\racets: Policy{-}Agnostic Programming in Racket

### 2018-07-23 paper/paper.tex
edits to intro

\newcommand{\comm}[3][\color{red}]{{#1{[{#2}: {#3}]}}}
\newcommand{\zw}[1]{\comm[\color{green!66!black}]{Zhanpeng}{#1}}
\newcommand{\kris}[1]{\comm[\color{blue}]{Kris}{#1}}
\newcommand{\tom}[1]{\comm[\color{orange}]{Tom}{#1}}
\newcommand{\lang}[0]{$\lambda_\textsc{FE}$\xspace}
\newcommand{\facet}[3]{{\fcol{\ensuremath{\langle \, {#1} \: ? \: {#2} \diamond \: {#3} \, \rangle}}}}
\newcommand{\formfacet}[3]{{\fcol{\ensuremath{\langle \langle \, {#1} \: ? \: {#2} \diamond \: {#3} \,  \rangle \rangle}}}}
\newcommand{\var}[1]{\mathit{#1}}

@@

\title{Racets: Policy{-}Agnostic Programming in Racket}

\author{Kristopher Micinski}
\affiliation{%
  \institution{Haverford College}
  \city{Haverford}
  \state{PA}
  \postcode{19041}
  \country{USA}}
\email{kris@cs.haverford.edu}
\author{Zhanpeng Wang}
\affiliation{%
  \institution{Haverford College}
  \city{Haverford}
  \state{PA}
  \postcode{19041}
  \country{USA}
}
\email{zwang10@haverford.edu}
\author{Thomas Gilray}
\affiliation{%
 \institution{University of Alabama}
 \city{Birmingham}
 \state{AL}
 \country{USA}}
\email{gilray@uab.edu}

@@

\keywords{security, faceted execution, information flow, macros}

@@

As our systems become more interconnected, they consume an
ever-growing amount of data.  System designers communicate to users
how their data is used via a \emph{privacy policy}. Unfortunately,
implementing these policies correctly is challenging: users often have
partial control over the policy (e.g., whether their phone number is
publicly visible or private) and policies change frequently (e.g.,
after new laws are passed). As policies evolve, developers face
massive (re)engineering efforts to ensure that implementations match
the espoused policy.

\emph{Policy-agnostic programming} is a linguistic paradigm that
decouples the implementation of privacy policies from the code that
operates on sensitive data. This frees developers to write programs as
they would normally, without inserting logic to manage the policy
directly into application code. Instead, data is labeled with its
policy as it enters the system, and policies propagate through the
program alongside data as computation progresses. Policy-agnostic
programming is often implemented using \emph{faceted execution}, a
dynamic information-flow monitor encoding sensitive values as
\emph{facets}: decision trees specifying different views of that
data. For example, the facet \facet{Alice}{\code{true}}{\code{false}}
represents a value that should appear to Alice as \blue{\code{true}} and to
everyone but Alice as \blue{\code{false}}.

Faceted execution propagates facets by extending core linguistic
primitives (such as function application) to handle faceted
values. For example, consider the application \blue{\code{(x true)}} where
\code{x} is \blue{\facet{Alice}{\lambda x.~\code{true}}{\code{not}}}.
Racket's standard function application will fail here because Racket's
\blue{\code{ppapp}} expects a procedure rather than a facet.  The correct way
to interpret function application on facets is to distribute the
application through each branch, producing
\facet{Alice}{\code{true}}{\code{false}}. Many other core forms (such
as \code{if}, \code{set!}, etc...) require similar changes.

Policy-agnostic programming promotes the idea that programmers should
be able to write programs ``normally'', without concerning themselves
with how privacy policies are enforced. And yet, we believe that the
relatively foundational linguistic changes required to enable faceted
execution have hindered implementations thusfar. For example, the
first implementation of faceted execution (by Austin et
al.~\cite{Austin:12}) extended a JavaScript interpreter to account for
faceted values. Other implementations use monads~\cite{Schmitz:2016}
or rely upon third-party macro systems~\cite{Yang:2016}. So far, we
know of no implementation of faceted execution that works seamlessly
alongside the language. By contrast, Racket has a powerful macro
system, allowing essentially any linguistic form (including function
application) to be extended with arbitrary semantics.

In this paper we present Racets, an implementation of policy-agnostic
programming in Racket via macros. Racets provides facilities for
creating policies and faceting secure data with those policies. Racets
also extends several core forms in Racket to work with faceted
values (our implementation is detailed in Section~\ref{sec:implementation}). We
have measured the overhead of Racets in several microbenchmarks, and
have used Racets to implement a small server-based
board-game (detailed in Section~\ref{sec:evaluation}). We see Racets as a
promising prototype for policy-agnostic programming in Racket, and
conclude with several future directions in Section~\ref{sec:conclusion}.

\section{Overview of Faceted Execution}

To introduce our setting we present the implementation of Battleship,
a small guessing-based board game, in Racets (this section presents a
distilled version of our case study in
Section~\ref{sec:case-study}). In this game each player has a
grid-based board on which they place tiles (or ``ships''). The players
hide their boards from each other as play progresses in rounds. Each
turn a player guesses the position of a tile on the other player's
board. If the guess is successful that tile is removed from the
board. Play ends once one player's board has no remaining tiles, at
which point that player loses.

@@

\section{Faceted Execution as Macros}

The semantics of faceted execution is an extension of the lambda
calculus, leading to a natural question: can we use Racket's macros to
extend Racket to faceted execution? We will see that the answer is
yes, and the translation from the big-step rules is surprisingly
straightforward. This section of our paper describes the design of
Racets, a prototype implementation of faceted execution using Racket
macros. In Section~\ref{sec:conclusion} we remark upon current
directions scaling Racets to the whole of Racket.

@@

Similarly, program counters are sets of branches. However, we must
still ask how we will keep track of the ``current'' program
counter. Our implementation uses Racket's parameters, though other
mechanisms (such as continuation marks\kris{cite}, to which parameters
macro-expand) could also be used. Racets defines the parameter
\code{current-pc}, and updates it as computation progresses:

@@

\begin{itemize}
\item Introduce an explicit \code{unbox} form in Racets, trusting the
  programmer to explicitly use our implementation of \code{unbox} on
  potentially-faceted objects.

\item Walk over Racket code (after macro-expansion via
  \code{local-expand}) transforming variable references to use
  explicit \code{unbox} forms from Racets.
\end{itemize}

For our prototype of Racets, we have chosen to implement the
first. This leads to a relatively simple implementation, but
essentially trusts the programmer to use Racets' \code{unbox} forms
when necessary.

\begin{figure}
\small
\begin{lstlisting}[language=Racket,escapeinside={<@}{@>},name=racets]
(define-syntax (ref-set! stx)
  (syntax-case stx ()
    [(_ var e)
     #`(let ([v e])
         (let write ([var var]
                     [pc (current-pc)])
           (if (box? var)
               ; <@$\color[rgb]{\colorCOMM}\var{write}(\sigma, \alpha, pc, v)$@>
               (set-box! var (construct-facet (current-pc) v (unbox var)))
               ; Else split
               (mkfacet
                 (facet-labelname (unbox var))
                 ; <@$\color[rgb]{\colorCOMM}\var{write}(\sigma, \alpha, pc \cup \{+l\}, v)$@>
                 (write
                   (facet-left (unbox var))
                   (set-add pc (pos (facet-labelname var))))
                 ; <@$\color[rgb]{\colorCOMM}\var{write}(\sigma, \alpha, pc \cup \{-l\}, v)$@>
                 (write
                   (facet-right (unbox var))
                   (set-add pc (neg (facet-labelname var))))))))]))
\end{lstlisting}
\caption{Racets' implementation of \code{set!}}
\label{fig:setbang}
\end{figure}

Racets defines a \code{box} macro, along with \code{unbox} and
\code{set!}. We include the definition of \code{set!} in
Figure~\ref{fig:setbang}, which inlines the definition of the
\textit{write} metafunction from Section~\ref{sec:semantics} to
consider the case under which a facet is used when an address is
expected. \zw{The role of the \textit{write} metafunction is to 
recursively dig into faceted value, and make writes to the addresses.
If it is an address at the beginning, which hits the base case,
the \textit{write} metafunction constructs a new faceted value according
to \bmth{pc} and makes modification to the address.}

@@

By now we can anticipate a predictable pattern for implementing
faceted execution: check \bmth{pc} to decide whether to branch left,
right, or split. This is largely our strategy for handling faceted
function application, with a small twist: we need to be able to apply
functions from outside of Racets. For example, if we want to apply
builtin functions such as \code{display}, we need to be mindful of the
fact that these functions cannot work with faceted arguments.

@@

In general it is unsafe to apply an unknown function to a faceted
value. This is because the unknown function may leak the facet's
private information as a side-effect. Therefore, our current
implementation of Racets allows programmers to apply external
functions, but does not make any guarantee of safety. A better
strategy may be to perform an \code{obs} before each call to a
potentially-unsafe function. In general, we believe module
interactions are a challenging problem in faceted execution, and we
leave its study to future work.

@@

We have implemented Racets as a set of Racket macros which can be
employed as a language using Racket's \texttt{\#lang reader}
facility. So far, we have included macros for many of Racket's core
forms including application, \code{if}, \code{lambda}, and
references. We leave others (including continuation marks) to future
work. Additionally, Racets does not support first-class control. There
has been recent work in handling exceptions in the context of faceted
execution~\cite{Austin:17}, however reconciling first-class control in
general remains (to our knowledge) an open problem.

@@

We plan to make our prototype of Racets available publicly.

\section{Conclusion}

>>>>

\newcommand{\comm}[3][\color{red}]{{#1{[{#2}: {#3}]}}}
\newcommand{\zw}[1]{\comm[\color{green!66!black}]{Zhanpeng}{#1}}
\newcommand{\kris}[1]{\comm[\color{blue}]{Kris}{#1}}
\newcommand{\tom}[1]{\comm[\color{orange}]{Tom}{#1}}
\newcommand{\lang}[0]{$\lambda_\textsc{FE}$\xspace}
\newcommand{\racets}[0]{$\textsc{Racets}$\xspace}
\newcommand{\facet}[3]{{\fcol{\ensuremath{\langle \, {#1} \: ? \: {#2} \diamond \: {#3} \, \rangle}}}}
\newcommand{\formfacet}[3]{{\fcol{\ensuremath{\langle \langle \, {#1} \: ? \: {#2} \diamond \: {#3} \,  \rangle \rangle}}}}
\newcommand{\var}[1]{\mathit{#1}}

@@

\title{\racets: Policy{-}Agnostic Programming in Racket}

\author{Kristopher Micinski}
\affiliation{%
  \institution{Haverford College}
  \city{Haverford}
  \state{PA}
  \postcode{19041}
  \country{USA}}
\email{kris@cs.haverford.edu}
\author{Zhanpeng Wang}
\affiliation{%
  \institution{Haverford College}
  \city{Haverford}
  \state{PA}
  \postcode{19041}
  \country{USA}
}
\email{zwang10@haverford.edu}
\author{Thomas Gilray}
\affiliation{%
 \institution{University of Alabama, Birmingham}
 \city{Birmingham}
 \state{AL}
 \country{USA}}
\email{gilray@uab.edu}

@@

\keywords{security, faceted execution, macros, information flow, languages as libraries}

@@

As information systems become more interconnected and complex, they consume an
ever-growing amount of private data. System designers communicate to users
how their data may be used via a \emph{privacy policy}. Unfortunately,
implementing such policies correctly is challenging: users often have
partial control over the policy (e.g., whether their phone number is
publicly visible or private) and policies can change frequently.
Not only can specific privacy policies be highly dynamic (dependent on runtime values),
but the process of improving privacy policies can be highly dynamic across time.
As policies evolve, developers face massive (re)engineering efforts to
ensure that implementations continue to match the policy at every relevant
point in the codebase.

\emph{Policy-agnostic programming} is a linguistic paradigm that
decouples the implementation of privacy policies from the code that
operates on sensitive data. This frees developers to write programs mostly
as they would for insecure code, without inserting specific logic to manage
information-flow policies directly into application code. Instead, data is
labeled with its policy as it enters the system and such labels propagate through
the program, alongside data, as computation progresses. When a secure value needs
to be introspected upon (or propagates outside the application), its policy
can be invoked at this point dynamically. This paradigm aims to permit
code manipulating sensitive data to be written in a manner entirely orthogonal to
policies themselves.

\emph{Faceted execution} (FE) is a highly expressive (but potentially expensive)
language semantics enabling policy-agnostic programming. In FE, dynamic information-flow
monitors instrument the program, encoding sensitive values as \emph{faceted values}:
decision trees specifying different views of data according to different possible security labels.
For example, the faceted value \facet{Alice}{\code{true}}{\code{false}}
represents a value that should appear to Alice as \blue{\code{true}} and to
everyone except Alice as \blue{\code{false}}. Faceted execution propagates distinct facets
of a value by extending core linguistic primitives (such as function application).
For example, consider the application \blue{\code{(x true)}} where
\code{x} is \blue{\facet{Alice}{\lambda x.~\code{true}}{\code{not}}}.
Racket's standard function application will fail here because Racket's
\blue{\code{ppapp}} expects a procedure rather than a facet. Instead, the proper way
to interpret function application on faceted values is to distribute the
application over all (in this case, both) facets, producing
\facet{Alice}{\code{true}}{\code{false}}: if Alice is viewing, the application yielded true,
otherwise it yielded false, so both are computed until the value is explicitly observed
with specified permissions. Many other core forms (such
as \code{if}, \code{set!}, etc...) require similar changes to handle faceted
values correctly.

Policy-agnostic programming promotes the idea that programmers should
be able to write programs ``normally'', without concerning themselves
with how privacy policies are enforced. Unfortunately, the
relatively foundational linguistic changes required to enable faceted
execution have hindered implementations thusfar. Dynamic generation
of first-class security labels, tracking an arbitrary number of
facets per value, and keeping facted-value trees in a canonical order,
are all central challenges in any practical implementation. For example, the
first implementation of FE (by \citet{Austin:2012}) extended
a JavaScript interpreter to account for faceted values. Other
implementations use monads~\cite{Schmitz:2016}
or rely upon third-party macro systems~\cite{Yang:2016}. We
know of no existing implementation of FE that aims to interoperate seamlessly
with code written in the host language. By contrast, Scheme boasts a powerful
hygenic macro system that allows essentially any linguistic form
to be modified arbitrarily.

In this paper we present \racets, an implementation of policy-agnostic
programming in Racket via macros. \racets provides facilities for
creating policies and faceting secure data with those policies. \racets
also extends several core forms in Racket to work with faceted
values (our implementation is detailed in Section~\ref{sec:implementation}). We
have measured the overhead of \racets in several microbenchmarks, and
have used \racets to implement a small server-based
board-game (detailed in Section~\ref{sec:overview}). We see \racets as a
promising prototype for policy-agnostic programming in Racket, and
conclude with discussion of future directions in Section~\ref{sec:conclusion}.

\section{Overview of Faceted Execution}
\label{sec:overview}

To introduce our setting we present the implementation of Battleship,
a small guessing-based board game, in \racets (this section presents a
distilled version of our case study in
Section~\ref{sec:case-study}). In this game each player has a
grid-based board on which they place tiles (or ``ships''). The players
hide their boards from each other as play progresses in rounds. Each
turn a player guesses the position of a tile on the other player's
board. If the guess is successful that tile is removed from the
board. Play ends once one player's board has no remaining tiles, at
which point that player loses.

@@

\section{Faceted Execution as Macros}
\label{sec:implementation}

The semantics of faceted execution is an extension of the lambda
calculus, leading to a natural question: can we use Racket's macros to
extend Racket to faceted execution? We will see that the answer is
yes, and the translation from the big-step rules is surprisingly
straightforward. This section of our paper describes the design of
\racets, a prototype implementation of faceted execution using Racket
macros. In Section~\ref{sec:conclusion} we remark upon current
directions scaling \racets to the whole of Racket.

@@

Similarly, program counters are sets of branches. However, we must
still ask how we will keep track of the ``current'' program
counter. Our implementation uses Racket's parameters, though other
mechanisms (such as continuation marks\kris{cite}, to which parameters
macro-expand) could also be used. \racets defines the parameter
\code{current-pc}, and updates it as computation progresses:

@@

\begin{itemize}
\item Introduce an explicit \code{unbox} form in \racets, trusting the
  programmer to explicitly use our implementation of \code{unbox} on
  potentially-faceted objects.

\item Walk over Racket code (after macro-expansion via
  \code{local-expand}) transforming variable references to use
  explicit \code{unbox} forms from \racets.
\end{itemize}

For our prototype of \racets, we have chosen to implement the
first. This leads to a relatively simple implementation, but
essentially trusts the programmer to use \racets' \code{unbox} forms
when necessary.

\begin{figure}
\small
\begin{lstlisting}[language=Racket,escapeinside={<@}{@>},name=racets]
(define-syntax (ref-set! stx)
  (syntax-case stx ()
    [(_ var e)
     #`(let ([v e])
         (let write ([var var]
                     [pc (current-pc)])
           (if (box? var)
               ; <@$\color[rgb]{\colorCOMM}\var{write}(\sigma, \alpha, pc, v)$@>
               (set-box! var (construct-facet (current-pc) v (unbox var)))
               ; Else split
               (mkfacet
                 (facet-labelname (unbox var))
                 ; <@$\color[rgb]{\colorCOMM}\var{write}(\sigma, \alpha, pc \cup \{+l\}, v)$@>
                 (write
                   (facet-left (unbox var))
                   (set-add pc (pos (facet-labelname var))))
                 ; <@$\color[rgb]{\colorCOMM}\var{write}(\sigma, \alpha, pc \cup \{-l\}, v)$@>
                 (write
                   (facet-right (unbox var))
                   (set-add pc (neg (facet-labelname var))))))))]))
\end{lstlisting}
\caption{\racets' implementation of \code{set!}}
\label{fig:setbang}
\end{figure}

\racets defines a \code{box} macro, along with \code{unbox} and
\code{set!}. We include the definition of \code{set!} in
Figure~\ref{fig:setbang}, which inlines the definition of the
\textit{write} metafunction from Section~\ref{sec:semantics} to
consider the case under which a facet is used when an address is
expected. \zw{The role of the \textit{write} metafunction is to 
recursively dig into faceted value, and make writes to the addresses.
If it is an address at the beginning, which hits the base case,
the \textit{write} metafunction constructs a new faceted value according
to \bmth{pc} and makes modification to the address.}

@@

By now we can anticipate a predictable pattern for implementing
faceted execution: check \bmth{pc} to decide whether to branch left,
right, or split. This is largely our strategy for handling faceted
function application, with a small twist: we need to be able to apply
functions from outside of \racets. For example, if we want to apply
builtin functions such as \code{display}, we need to be mindful of the
fact that these functions cannot work with faceted arguments.

@@

In general it is unsafe to apply an unknown function to a faceted
value. This is because the unknown function may leak the facet's
private information as a side-effect. Therefore, our current
implementation of \racets allows programmers to apply external
functions, but does not make any guarantee of safety. A better
strategy may be to perform an \code{obs} before each call to a
potentially-unsafe function. In general, we believe module
interactions are a challenging problem in faceted execution, and we
leave its study to future work.

@@

We have implemented \racets as a set of Racket macros which can be
employed as a language using Racket's \texttt{\#lang reader}
facility. So far, we have included macros for many of Racket's core
forms including application, \code{if}, \code{lambda}, and
references. We leave others (including continuation marks) to future
work. Additionally, \racets does not support first-class control. There
has been recent work in handling exceptions in the context of faceted
execution~\cite{Austin:17}, however reconciling first-class control in
general remains (to our knowledge) an open problem.

@@

We plan to make our prototype of \racets available publicly.

\section{Conclusion}
\label{sec:conclusion}

asdf asdf
